Hierarchical and grouped time series forecasting: an application to Emergency Medical Services

### 2022-10-27 manuscript/main.qmd

---
title: "Hierarchical and grouped time series forecasting: an application to Emergency Medical Services"
execute:
  echo: false
  message: false
  warning: false
  cache: true
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Annals of Emergency Medicine
      formatting: preprint
      model: 3p
      cite-style: authoryear
  elsevier-html: 
    toc: true
author:
  - name: Bahman Rostami-Tabar
    email: rostami-tabarb@cardiff.ac.uk
    affiliations: 
        - id: carbs
          name: Cardiff University
          department: Cardiff Business School
          country: United Kingdom
          postal-code: CF10 3EU
    attributes:
        corresponding: true
  - name: Rob J. Hyndman
    email: Rob.Hyndman@monash.edu
    affiliations:
        - id: monash
          name: Monash University
          department: Department of Econometrics and Business Statistics
          country: Australia
          postal-code: VIC 3800
abstract: |
  The primary goal of an emergency service is to minimize the ambulance response time. Accurate forecasts of ambulance demand are crucial inputs into the decisions of planners and policymakers on the resource (e.g. staff and fleet) planning and dynamic deployment. Informing those decisions require detailed forecasting of demand at multiple levels of hierarchy from the national and regional levels to the nature of incidents and their priorities. These forecasts are often generated independently by different teams. However, independent forecasts may not add up to the other levels resulting in mismatched forecasts that lead to conflicting decisions and a lack of coordination in the care service. To address this issue, we explore the inherently hierarchical and grouped structure of Emergency Medcine Services. We apply forecast reconciliation methods to generate both point and probabilistic forecasts that are coherent, for the first time in the healthcare literature.
keywords: [Emergency services, healthcare, hierarchical time series forecasting, forecast reconciliation]
bibliography: mybibfile.bib
---

>>>>

---
title: "Hierarchical and grouped time series forecasting: an application to Emergency Medical Services"
author:
  - name: Bahman Rostami-Tabar
    email: rostami-tabarb@cardiff.ac.uk
    affiliations: 
        - id: carbs
          name: Cardiff University
          department: Cardiff Business School
          country: United Kingdom
          postal-code: CF10 3EU
    attributes:
        corresponding: true
  - name: Rob J. Hyndman
    email: Rob.Hyndman@monash.edu
    affiliations:
        - id: monash
          name: Monash University
          department: Department of Econometrics and Business Statistics
          country: Australia
          postal-code: VIC 3800
abstract: |
  The primary goal of an emergency service is to minimize the ambulance response time. Accurate forecasts of ambulance demand are crucial inputs into the decisions of planners and policymakers on the resource (e.g. staff and fleet) planning and dynamic deployment. Informing those decisions require detailed forecasting of demand at multiple levels of hierarchy from the national and regional levels to the nature of incidents and their priorities. These forecasts are often generated independently by different teams. However, independent forecasts may not add up to the other levels resulting in mismatched forecasts that lead to conflicting decisions and a lack of coordination in the care service. To address this issue, we explore the inherently hierarchical and grouped structure of Emergency Medcine Services. We apply forecast reconciliation methods to generate both point and probabilistic forecasts that are coherent, for the first time in the healthcare literature.
keywords: 
  - Healthcare
  - Emergency services
  - forecast reconciliation
  - hierarchical time series forecasting
bibliography: bibliography.bib
execute:
  echo: false
  message: false
  warning: false
  cache: true
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Annals of Emergency Medicine
      formatting: preprint
      model: 3p
      cite-style: authoryear
  elsevier-html: 
    toc: true
---
